---
title: "R Notebook"
output: html_notebook
---

```{r}
install.packages("tidyverse")
```


```{r}

beer <- read_excel("Zakupy.xls")
names(beer) <- tolower(names(beer)) #lowercase for names
head(beer) #display

```
```{r}
ggplot(data = beer, aes(x = wiek, color = factor(plec))) + geom_density() + facet_wrap(~towar)
```
```{r}
ggplot(data = beer, aes(x = wiek, color = factor(towar))) + geom_density() + facet_wrap(~plec)
```
### Making a logistic regression model
```{r}
m1 <- glm(formula = towar ~ wiek + plec, data = beer, family = binomial)

summary(m1)
```

## Odds ratio
```{r}
m1 |> coef() |> exp() |> round(4)
```

## Compare women in their 40s with women in their 30s
```{r}
exp(-0.18668*40 - - 0.18668*30)
```

## Plot of regression model
```{r}
curve(exp(7.16164 -0.18668*x + 0.96478)/(1+exp(7.16164 -0.18668*x + 0.96478)), from = 18, to = 60, xlab = "Age", y = "P(Towar=1)")
curve(exp(7.16164 -0.18668*x)/(1+exp(7.16164 -0.18668*x)), from = 18, to = 60, xlab = "Age", y = "P(Towar=1)", add = T, col = "red")
```

## Pseudo R^2
```{r}
1 - m1$deviance / m1$null.deviance
```

## Probabilities
Probability that someone will buy a beer of given type
```{r}
prob1 <- predict(m1, type = "response")
head(prob1)
```

## Confusion matrix
Add thresholds for bought or not bought a beer.
```{r}
beer$prob <- prob1
```

```{r}
xtabs(~ I(prob>=0.65)+towar, data = beer) |> prop.table(margin = 1)
```
```{r}
roc_data <- calculate_roc(beer$prob, beer$towar)
head(roc_data)
ggroc(roc_data)
```
Top left corner (around 0.65 threshold) is the best here! We are only minimizing the errors of last row (false positives) though.
